---
title: "GSS challange Less Data"
author: "Eliana Martinez"
date: "spring 25"
output: html_document
---



```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,sex,race,cappun,degree,relig,gunlaw,tvhours,owngun,age,wordsum)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd3 <- mydata %>% group_by(cappun,race) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
 
 
 # the group_by followed by summarize(count=n())
 
basicC <- ggplot(dd3,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")
 
 #Now for percentage plot
 
basicCC <- ggplot(dd3,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The data reveals distinct mathematical disparities in attitudes toward capital punishment across racial groups. Black respondents show the strongest opposition, with opposition counts far exceeding support. White respondents display the highest favorability, with support clearly outpacing opposition. The "other" group trends toward balance but still leans slightly toward opposition. The most significant contrasts appear between white and Black respondents—whites strongly in favor, Blacks strongly opposed—while the "other" group reflects a smaller yet consistent tilt toward opposition. These patterns underscore measurable differences in public sentiment by race.

## Now with LESS data (more variables give less data due to NA's) create the approate graphic displays to answer the question.





### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table2 <- xtabs(~cappun + race, data=mydata)
```

The data shows the white group favors capital punishment with a ratio of about 2:1, meaning for every two people in favor, one is opposed. The black group is more evenly split but slightly against, suggesting their opposition is marginally higher than support. The "other" group opposes capital punishment more strongly, with a ratio of around 3:2, indicating for every three people opposed, two favor it. These ratios highlight differing levels of support and opposition across racial groups.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

`
```{r}
chisq.test(table2)
```
 
The p-value of 0.00005896 indicates a highly significant association between the variables, suggesting the observed differences are unlikely due to random chance. This strong result warrants further analysis. Conducting a Fisher’s exact test will help assess the statistical significance and relationships between groups with greater precision, offering a more detailed understanding of the patterns within the data.

```{r}
 fisherXY(mydata$race, mydata$cappun, confL = 0.99, printAll = TRUE)
```
The p-values for the comparisons reveal distinct relationships between the racial groups. The p-value for the comparison between white and black respondents is 0.00001741, indicating a highly significant difference, suggesting a strong association in their opinions. In contrast, the comparison between white and other respondents yields a p-value of 0.2565, which is not statistically significant, indicating no meaningful difference in their views. Similarly, the p-value for the black and other group comparison is 0.06299, approaching significance but still above the conventional threshold. Overall, these p-values illustrate varying degrees of difference among the groups, highlighting significant disparities in opinions on capital punishment.

# Discussion and Conclusion

The number 2 branch analysis, based on 821 variables, shows that white and "other" groups generally support capital punishment, while the Black group predominantly opposes it. This contrasts with the master branch, which uses 3,000 variables and yields markedly different visual and statistical outcomes. In the number 2 branch, white respondents show a 2:1 support ratio, while the master branch reflects a narrower 2:3 ratio. The “other” group supports capital punishment in the number 2 branch but leans more toward opposition in the master branch. Meanwhile, the Black group’s opposition shifts from a strong 2:1 ratio in the master branch to a 2:3 ratio in the number 2 branch. These differences underscore how variable selection can significantly shape the interpretation of public sentiment across racial groups.
<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
