---
title: "GSS challange More Data"
author: "Eliana Martinez"
date: "Spring 2025"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
dd3 <- mydata %>% group_by(race,cappun) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd3,aes(x=race,y=count,fill=cappun))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd3,aes(x=race,y=prcnt*100,fill=cappun)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
From our observations, we can see a dark difference in how white people vote in comparison to other races about 1200 white people voted in favor of punishment while less than 250 Black people and other races voted in favor of that we can also see that in general, whether in favor or against capital punishment a lot more white people voted.

## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

 
```{r}
table1 <- xtabs(~race + cappun, data=mydata)
```
The data shows opinions on capital punishment (cappun) segmented by race (white, black, and other). Among white respondents, 79.80% favor capital punishment, while only 11.72% oppose it. In contrast, black respondents show significantly lower support, with 65.38% favoring and 24.11% opposing. For those categorized as "other," 8.47% favor capital punishment. The data suggests a clear disparity: white respondents predominantly favor capital punishment, while black respondents display more opposition. This trend indicates varying racial pe
## Now with More data



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
fisherXY(mydata$race, mydata$cappun, confL = 0.99 )
```

In the Fisher's exact test, all p-values were low, indicating significant associations between racial groups and opinions on capital punishment. The p-value for white and black respondents was 0.00000000000000022, demonstrating a highly significant difference. For white and other respondents, the p-value was 0.002607, also indicating a significant association. Additionally, the p-value for black and other respondents was 0.001564, further supporting the presence of a statistically significant relationship. Overall, these results highlight the differing opinions on capital punishment among racial groups, underscoring the importance of examining these disparities.

## More data version


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
Analysis from the master branch, using 3,000 variables, shows that the white group largely supports capital punishment, while Black and "other" groups strongly oppose it. In contrast, the number 2 branch, based on a narrower dataset of 821 variables, presents different trends: the white group shows a 2:1 support ratio, compared to a closer 2:3 ratio in the master branch. The “other” group is more supportive in the number 2 branch but more opposed in the master branch. Similarly, Black respondents show a 2:1 opposition in the master branch, shifting to 2:3 in the number 2 branch. These variations highlight how dataset scope and variable selection can significantly shape interpretations of public opinion and influence policy debates on capital punishment. 
